---
title: "Racism against Black and Asian American populations in COVID-19"
author: "Vivacious Vaxxers"
date: "May 2021"
output:
  rmdformats::readthedown:
    thumbnails: false
    highlight: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
#loading libraries
library(tidyverse)
library(DT)
library(mosaic)
library(foreign)
#loading datasets
#police shootings
acab <- read_csv("data/fatal-police-shootings-data.csv") %>%
  filter(date > "2019-12-31")
#hate crimes
hateCrimes <-  read_csv("data/table-1.csv") %>%
	janitor::clean_names() %>%
	select(bias_motivation, incidents)
#us population stats
pop <- read_csv("data/pop_stats.csv") %>%
	janitor::clean_names() %>%
	na.omit()
```

# Motivations
  
For our final project, we will not be building upon our mid-semester project, but instead looking at the social impacts of the pandemic specifically on Asian- and Black-American populations. Through our mid semester project we realized that the current available data surrounding vaccinations isn't all that great; however, we still wanted to focus on COVID and the sociopolitical effects of the pandemic which have a multitude of public health implications. Additionally, this past year has brought the prevalence of anti-Black and anti-Asian hate crimes and violence to the forefront of many American minds. This issue is not just a social one but greatly impacts health, and as of early April, [the CDC has declared racism](https://www.cdc.gov/media/releases/2021/s0408-racism-health.html) as a threat to public health.

# Research Questions
By analyzing both demographic, health, and social data (such as social media metrics), our group aims to explore how the past pandemic year has impacted both the mental and physical health of our chosen population. Our guiding questions are:
1. How has the pandemic impacted access to  healthcare/social services?
2. How has the pandemic impacted mental health?
3. Have popular social movements such as [#StopAsianHate](https://twitter.com/search?q=%23StopAsianHate&src=typeahead_click) or [#BlackLivesMatter](https://twitter.com/search?q=%23StopAsianHate&src=typeahead_click) had an impact on health outcomes?

# Why focus on COVID-19



# Communities of Interest

## Black America


## Asian America

```{r echo=FALSE, warning=FALSE, message=FALSE}
#hate crimes bar chart
g <- hateCrimes %>%
	ggplot(aes(x="", y=incidents, fill=bias_motivation))
g + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
	theme_void() +
	labs(x= "Bias Motivation",
			 y= "Number of Incidents",
			 title = "Number of bias related incidents",
			 subtitle = "in 2019",
			 fill= NULL)
#police shootings pie chart
#g2 <- acab %>%
	#ggplot(aes(x="", y=race, fill=race)) 
  
#g2 + geom_bar(stat="identity", width=1) +
  #coord_polar("y", start=0) +
#	theme_void() +
  #labs(title="Number of Police Killings by race",
       #subtitle = "2020 to 2021",
			# caption = "A=Asian, B=Black, H=Hispanic,\n N=Native, O=Other, W=White",
			  #fill= "Race")
#population pie chart
g3 <- pop %>%
	ggplot(aes(x="", y=united_states, fill=fact))
g3 + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
	theme_void() +
	labs(title= "Percentage of US racial groups",
			 subtitle = "from 2020 census",
			 fill="Racial group")
```

Check out some of our spatial visualizations [here](https://lewis-ayodele.shinyapps.io/BlogAL/)

# Twitter Analytics

# Mental Health



You should test out updating your GitHub Pages website:

- clone your group's blog project repo in RStudio
- update "Your Project Title Here" to a new title in the YAML header
- knit `index.Rmd` 
- commit and push BOTH the `index.Rmd` and the `index.html` files
- go to https://stat231-s21.github.io/Blog-VIVACIOUS-VAXXERS/ to see the published test document (this is publicly available!)


