---
title: "Racism against Black and Asian American populations in COVID-19"
author: "Vivacious Vaxxers"
date: "May 2021"
output:
  rmdformats::readthedown:
    thumbnails: false
    highlight: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
#loading libraries
library(tidyverse)
library(DT)
library(mosaic)
library(foreign)
#loading datasets
#police shootings
acab <- read_csv("data/fatal-police-shootings-data.csv") %>%
  filter(date > "2019-12-31")
#hate crimes
hateCrimes <-  read_csv("data/table-1.csv") %>%
	janitor::clean_names() %>%
	select(bias_motivation, incidents)
#us population stats
pop <- read_csv("data/pop_stats.csv") %>%
	janitor::clean_names() %>%
	na.omit()
```

# Motivations
  
For our final project, we will not be building upon our mid-semester project, but instead looking at the social impacts of the pandemic specifically on Asian- and Black-American populations. Through our mid semester project we realized that the current available data surrounding vaccinations isn't all that great; however, we still wanted to focus on COVID and the sociopolitical effects of the pandemic which have a multitude of public health implications. Additionally, this past year has brought the prevalence of anti-Black and anti-Asian hate crimes and violence to the forefront of many American minds. This issue is not just a social one but greatly impacts health, and as of early April, [the CDC has declared racism](https://www.cdc.gov/media/releases/2021/s0408-racism-health.html) as a threat to public health.

# Research Questions
By analyzing both demographic, health, and social data (such as social media metrics), our group aims to explore how the past pandemic year has impacted both the mental and physical health of our chosen population. Our guiding questions are:
1. How has the pandemic impacted access to  healthcare/social services?
2. How has the pandemic impacted mental health?
3. Have popular social movements such as [#StopAsianHate](https://twitter.com/search?q=%23StopAsianHate&src=typeahead_click) or [#BlackLivesMatter](https://twitter.com/search?q=%23StopAsianHate&src=typeahead_click) had an impact on health outcomes?

# Why focus on COVID-19

Our group wanted to focus on Black and Asian communities during the pandemic because these are the communities that hit closest to home for our research group. When the pandemic first started with the president of the United States calling the virus, [the Chinese virus](https://www.ucsf.edu/news/2021/03/420081/trumps-chinese-virus-tweet-linked-rise-anti-asian-hashtags-twitter), the killings of [George Floyd](https://en.wikipedia.org/wiki/George_Floyd) and [Breonna Taylor](https://en.wikipedia.org/wiki/Killing_of_Breonna_Taylor) occurring in the summer of 2020, and the disproportionate deaths of people of color from [COVID-19](https://www.cdc.gov/coronavirus/2019-ncov/covid-data/investigations-discovery/hospitalization-death-by-race-ethnicity.html), we have personally perceived changes in our lived experiences as Black and Asian Americans. From our perspective, racism and bias has only seemed to increase during this pandemic despite growing awareness of the plights our communities face. We wanted to take a data based approach to see if our lived experiences and intuitions were reflected in publicly available national data. 

# Communities of Interest

While our communities of interest, Black and Asian America, are in no ways monolithic, there are data that we can focus on of these communities pre-COVID to get a better understanding of how the average experience has changed over the COVID-19 pandemic. 

## Black America

Police brutality against the Black community has come to the forefront of national conciousness over the past year with highly publicized murders by police such as the murder of George Floyd and Breonna Taylor. However, violence at the hands of the state is nothing new for this population. Black men and women are significantly more likely to be killed by police than their white counterparts [(Edwards et al 2019)](https://www.pnas.org/content/116/34/16793). Additionally, Black people disproportionately face hate crime and bias related incidents (see pie chart below). 
But the story Black America is not just one of plight. There is growing diversity within the Black population and an overall population increase. [Pew reports](https://www.pewresearch.org/social-trends/2021/03/25/the-growing-diversity-of-black-america/) that 1 in 4 Black people are members of generation Z, and there is a steady increase of Black people receiving college degrees.


## Asian America

The COVID-19 pandemic continues to spotlight the increase in anti-Asian bias and hate across the country. But, anti-Asian bias has been a crisis in this country for centuries. Laws such as the [Chinese Exclusion Act](https://history.state.gov/milestones/1866-1898/chinese-immigration) and [Japanese internment](https://www.archives.gov/education/lessons/japanese-relocation) codified anti-Asian sentiment. In April 2021, Pew Research reported that [1 in 3 Asian Americans fear race based physical violence](https://www.pewresearch.org/fact-tank/2021/04/21/one-third-of-asian-americans-fear-threats-physical-attacks-and-most-say-violence-against-them-is-rising/). 
But like Black-America, the narrative is not only negative. Asian-Americans are the fastest growing racial group in America, and they are more likely to have a college degree than the average American. Overall, Asian-Americans have lower poverty rates than the national average, but withingroup differences are large, and smaller impoverished Asian ethnic groups are often overlooked  [(Pew 2021)](https://www.pewresearch.org/fact-tank/2021/04/09/asian-americans-are-the-fastest-growing-racial-or-ethnic-group-in-the-u-s/). It is important to note that it is hard to come by strong demographic data for asian populations in the US because many national surveys, such as the surveys conducted by Pew Research center, are not conducted in Asian languages, therefore there is a strong non-response bias in our data. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
#hate crimes bar chart
g <- hateCrimes %>%
	ggplot(aes(x="", y=incidents, fill=bias_motivation))
g + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
	theme_void() +
	labs(x= "Bias Motivation",
			 y= "Number of Incidents",
			 title = "Number of bias related incidents",
			 subtitle = "in 2019",
			 fill= NULL)
#police shootings pie chart
#g2 <- acab %>%
	#ggplot(aes(x="", y=race, fill=race)) 
  
#g2 + geom_bar(stat="identity", width=1) +
  #coord_polar("y", start=0) +
#	theme_void() +
  #labs(title="Number of Police Killings by race",
       #subtitle = "2020 to 2021",
			# caption = "A=Asian, B=Black, H=Hispanic,\n N=Native, O=Other, W=White",
			  #fill= "Race")
#population pie chart
g3 <- pop %>%
	ggplot(aes(x="", y=united_states, fill=fact))
g3 + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
	theme_void() +
	labs(title= "Percentage of US racial groups",
			 subtitle = "from 2020 census",
			 fill="Racial group")
```

Check out some of our spatial visualizations [here](https://lewis-ayodele.shinyapps.io/BlogAL/)

# Twitter Analytics

# Mental Health

[Click here](https://lewis-ayodele.shinyapps.io/MentalHealthBK/) to see our interactive visualizations!




